---
title: "ANALYTE_NAME_HERE exchange scheme" # change this to your analyte of choice
header-includes:
  \usepackage{palatino}
  \renewcommand{\familydefault}{\sfdefault}
  \fontfamily{ppl}\selectfont
  \usepackage{booktabs}
  \usepackage{longtable}
output:
  pdf_document:
    df_print: kable
  html_document: default
fontsize: 12pt # you can alter the master font size here
---

<!-- 
READ THIS FIRST!

Note that you will need to install rmarkdown on your machine prior to running this document. This comes pre-packaged with RStudio, but if you're using something else, then this can be installed with the following command in the R console:

install.packages("rmarkdown")

You will need also need to install TinyTex - which is what generates the PDF behind the scenes - with the R console commands below:

install.packages("tinytex")
tinytex::install_tinytex()

This make take several minutes to complete. It is also worth checking that it is all set up correctly by using the following in the R console:

tinytex::is_tinytex()

This should output the result "TRUE" in the console.

-->

***

```{r load_packages, echo=FALSE, message=FALSE, warning=FALSE}
# These are the required R packages. RStudio will automatically check to see if they're installed
library(DescTools)
library(dplyr)
library(ggplot2)
library(irr)
library(kableExtra)
library(knitr)
library(reshape2)
options(tinytex.verbose = TRUE)

# Plot theme settings - change for your preference!
theme_plot <- theme(
  panel.background = element_blank()
  ,panel.grid.major = element_blank()
  ,panel.grid.minor = element_blank()
  ,panel.border = element_rect(fill = NA, colour = "black")
  ,axis.text = element_text(colour = "black", size = 12)
  ,axis.title = element_text(colour = "black", size = 12)
  ,legend.text = element_text(colour = "black", size = 12)
  ,legend.title = element_text(colour = "black", size = 12, face = "bold")
  ,legend.background = element_rect(colour = "black", fill = NA)
  ,axis.ticks = element_line(colour = "black")
  ,strip.background = element_rect(fill = "grey80", colour = "black")
  ,strip.text = element_text(colour = "black", size = 12)
)

# Global variables
num_digits <- 1 # number of digits to round values to
num_samples <- 3 # number of samples per distribution
window_size <- 6 # number of distributions to calculate B- and C-scores from
units <- "nmol/L"
analyte_name <- "YOUR_ANALYTE_HERE"
author_name <- "YOUR_NAME_HERE"
file_path <- "YOUR_FILE_PATH_HERE"
```

```{r read_data, echo=FALSE, message=FALSE, warning=FALSE}
# Read in csv file of raw data
df <- read.csv(
  # change this to the file path of your input data, although I've sent you some example data
  file = file_path
  ,header = TRUE  # this states that the first row is the column names
)

# Let's say that this is the structure of your data, where each distribution has three samples
# Note that this is in long format already for simplicity

# ------------------------------------
# Distribution | Sample | Lab | Result
# ------------------------------------
#            1 |      1 |   A |    1.4
#            1 |      1 |   B |    2.2
#            1 |      1 |   C |    1.6
#            1 |      2 |   A |    3.6
#            1 |      2 |   B |    4.2
#            1 |      2 |   C |    3.9
#            1 |      3 |   A |   10.1
#            1 |      3 |   B |   11.3
#            1 |      3 |   C |   10.5
#            2 |      1 |   A |    5.4
#            2 |      1 |   B |    6.5
#            2 |      1 |   C |    6.1
#            2 |      2 |   A |    1.6
#            2 |      2 |   B |    2.5
#            2 |      2 |   C |    1.7
#            2 |      3 |   A |   22.3
#            2 |      3 |   B |   21.1
#            2 |      3 |   C |   24.2
# ------------------------------------

# Get last distribution number
num_dist <- max(unique(as.numeric(as.character(df$Distribution))))
```
### **Report author:**

`r author_name`

### **Report date:**

`r date()`

### **Distribution number:**

`r num_dist`

### **Summary of distribution `r num_dist`'s results**:
```{r show_data, echo=FALSE, message=FALSE, warning=FALSE}
df_results <- filter(df, Distribution == num_dist)

kable(df_results, booktabs = TRUE, align = c("r", "r", "l", "r")) %>%
  kable_styling(position = "left", font_size = 12) %>%
  # NB: you will need to change the group_rows commands here depending on how many samples you distribute
  group_rows("1", 1, num_samples) %>%
  group_rows("2", num_samples + 1, 2 * num_samples) %>%
  group_rows("3", (2 * num_samples) + 1, 3 * num_samples)
```
\clearpage

### **Absolute difference vs ALM (all-laboratory mean)**:
```{r plot_diff_vs_alm, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=6, fig.align='center'}
# Calculate ALM, ALSD, bias, and difference for each distribution
df_processed <- df %>%
  group_by(Distribution, Sample) %>%
  mutate(
    alm = round(mean(Result, na.rm = TRUE), num_digits)
    ,alsd = sd(Result, na.rm = TRUE)
    ,difference = round(Result - alm, num_digits)
    ,bias = round(difference / alm, 3) * 100
  ) 

# Individual differences ~ Concentration
max_diff = max(filter(df_processed, Distribution == num_dist)$difference, na.rm = TRUE)
min_diff = min(filter(df_processed, Distribution == num_dist)$difference, na.rm = TRUE)

if (abs(max_diff) > abs(min_diff)) {
  y_max <- max_diff
  y_min <- -max_diff
} else {
  y_max <- abs(min_diff)
  y_min <- min_diff
}

# Plot 1
plot_1 <- ggplot(filter(df_processed, Distribution == num_dist), aes(x = alm, y = difference, colour = Lab))+
  geom_point(size = 3, alpha = 0.75)+
  geom_hline(yintercept = 0, colour = "black", linetype = "dashed")+
  theme_plot+
  expand_limits(x = 0)+
  xlab(paste0("ALM (", units, ")"))+
  ylab(paste0("Absolute difference ", units))+
  ylim(c(y_min, y_max))
print(plot_1)

```

### **Relative bias vs ALM**:
```{r plot_bias_vs_alm, echo=FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=6, fig.align='center'}
# Bias ~ Concentration
max_bias = max(filter(df_processed, Distribution == num_dist)$bias, na.rm = TRUE)
min_bias = min(filter(df_processed, Distribution == num_dist)$bias, na.rm = TRUE)

if (abs(max_bias) > abs(min_bias)) {
  y_max <- max_bias
  y_min <- -max_bias
} else {
  y_max <- abs(min_bias)
  y_min <- min_bias
}

plot_2 <- ggplot(filter(df_processed, Distribution == num_dist), aes(x = alm, y = bias, colour = Lab))+
  geom_hline(yintercept = 0, colour = "black", linetype = "dashed")+
  geom_point(size = 3, alpha = 0.75)+
  ylim(c(y_min, y_max))+
  expand_limits(x = 0)+
  theme_plot+
  xlab(paste0("ALM (", units, ")"))+
  ylab("Relative bias (%)")
print(plot_2)
```
\clearpage

### **Relative bias vs distribution**:
```{r plot_bias_vs_dist, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5.33, fig.height=8.6, fig.align='center'}
# Total distributions
distributions <- as.numeric(unique(df_processed$Distribution))

plot_3 <- ggplot(df_processed, aes(x = as.numeric(Distribution), y = bias))+
  geom_hline(yintercept = 0, linetype = "dashed")+
  geom_point(size = 3, alpha = 0.5)+
  facet_wrap(~Lab, ncol = 1)+
  theme_plot+
  ylim(c(-50, 50))+
  scale_x_continuous(breaks = distributions)+
  xlab("Distribution")+
  ylab("Relative bias (%)")
print(plot_3)
```
\clearpage

### **Relative bias vs concentration**:
```{r plot_bias_vs_conc, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5.33, fig.height=8.6, fig.align='center'}
plot_4 <- ggplot(df_processed, aes(x = Result, y = bias))+
  geom_hline(yintercept = 0, linetype = "dashed")+
  geom_point(size = 3, alpha = 0.6)+
  facet_wrap(~Lab, ncol = 1)+
  theme_plot+
  ylim(c(-50, 50))+
  xlab(paste0(analyte_name, " (", units, ")"))+
  ylab("Relative bias (%)")
print(plot_4)
```
\clearpage

### **B-score summary**:
B-scores are based on the mean of the last `r window_size` distributions' biases (in a rolling window). The number of data-points used to calculate the B-score for each distribution is shown above each point. Data-points may be missing due to undetectable ("<") values.

```{r b_scores, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7.5, fig.width=5.33, fig.align='center'}
# Generate vectors of `window_size` distribution windows (trim for values < 1) and calculate B & C scores
list_windows <- list()

for (i in rev(distributions)) {
  
  dist_window <- (i - (window_size - 1)):i
  dist_window <- dist_window[which(dist_window > 0)]
  dist_window <- as.character(dist_window)
  
  df_summary <- df_processed %>%
    filter(Distribution %in% dist_window) %>%
    group_by(Lab) %>%
    summarise(
      n = sum(!is.na(bias))
      ,B_score = ifelse(
        n >= 3 
        ,yes = mean(bias, na.rm = TRUE)
        ,no = NA
      )
      ,C_score = ifelse(
        n >= 3 
        ,yes = sd(bias, na.rm = TRUE)
        ,no = NA
      )
    )
  df_summary$Distribution <- i
  df_summary$Window <- paste0(dist_window, collapse = "")
  list_windows[[i]] <- df_summary
  
}

df_summary_all <- bind_rows(list_windows)

# Convert NaN to NA
df_summary_all$B_score[is.na(df_summary_all$B_score)] <- NA 
df_summary_all$C_score[is.na(df_summary_all$C_score)] <- NA 

# Calculate scales
max_b = max(df_summary_all$B_score, na.rm = TRUE)
min_b = min(df_summary_all$B_score, na.rm = TRUE)

if (abs(max_b) > abs(min_b)) {
  y_max <- max_b + (0.4 * max_b)
  y_min <- -(max_b + (0.4 * max_b))
} else {
  y_max <- abs(min_b + (0.4 * min_b))
  y_min <- min_b + (0.4 * min_b)
}

# B score plots
plot_5 <- ggplot(df_summary_all, aes(x = as.numeric(Distribution), y = B_score))+
  geom_hline(yintercept = 0, linetype = "dashed")+
  geom_point(size = 3)+
  geom_line(aes(group = Lab))+
  geom_text(aes(label = n, y = B_score + 8))+
  facet_wrap(~Lab, ncol = 1)+
  theme_plot+
  scale_x_continuous(breaks = distributions)+
  scale_y_continuous(limits = c(y_min, y_max), expand = c(0.2, 0))+
  xlab("Distribution")+
  ylab("B score (%)")
print(plot_5)

```
\clearpage

### **C-score summary**:
C-scores are based on the SD of the last `r window_size` distributions' biases (in a rolling window). The number of data-points used for each calculated C-score are as shown in the B-score plot on page 7.

```{r c_scores, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7.5, fig.width=5.33, fig.align='center'}
# C score plots
plot_6 <- ggplot(df_summary_all, aes(x = as.numeric(Distribution), y = C_score))+
  geom_point(size = 3)+
  geom_hline(yintercept = 0, linetype = "dashed")+
  geom_line(aes(group = Lab))+
  facet_wrap(~Lab, ncol = 1)+
  theme_plot+
  scale_x_continuous(breaks = distributions)+
  ylim(c(0,30))+
  xlab("Distribution")+
  ylab("C score (%)")
print(plot_6)
```
<!-- \clearpage -->

### **Tabulated B- and C- score data**:
```{r summary_table, echo=FALSE, warning=FALSE, message=FALSE}
dist_range <- tail(distributions, window_size)

df_scores <- df_summary_all %>%
  arrange(Lab) %>%
  select(Lab, Distribution, B_score, C_score) %>%
  mutate(B_score = format(B_score, digits = 2), C_score = format(C_score, digits = 2)) %>%
  rename(`B-score` = B_score, `C-score` = C_score)

filter(df_scores, Distribution %in% dist_range) %>%
  kable(format = "latex", booktabs = TRUE, align = c("r", "l", "r", "r")) %>%
  kable_styling(position = "left", font_size = 12) %>%
  # NB: you will need to change the group_rows commands here depending on how many labs you have participating
  group_rows("A", 1, length(dist_range)) %>%
  group_rows("B", length(dist_range) + 1, 2 * length(dist_range)) %>%
  group_rows("C", (2 * length(dist_range)) + 1, 3 * length(dist_range))
```
\clearpage
